---
title: "R coding Exercise"
output: 
  html_document:
    toc: FALSE
---
```{r}
#load dslabs package
library("dslabs")
library("ggplot2")
library("dplyr")
library("broom")
```

```{r}
#look at help file for gapminder data
help(gapminder)
#get an overview of data structure
str(gapminder)
#get a summary of data
summary(gapminder)
#determine the type of object gapminder is
class(gapminder)
```
```{r}
#create new object selecting only African countries
africadata <-  gapminder[which(gapminder$continent=='Africa'), ]
#overview of the new data structure
str(africadata)
summary(africadata)
```

```{r}
#making two variable from new dataset
inmor_liexp <- africadata[c("infant_mortality","life_expectancy")]
pop_liexp <- africadata[c("population","life_expectancy")]
#overview and summary of both
str(inmor_liexp)
str(pop_liexp)
summary(inmor_liexp)
summary(pop_liexp)
```

```{r}
#plotting life expectancy as a function of infant mortality [have to install ggplot2 first]
ggplot(inmor_liexp, aes(x= infant_mortality, y = life_expectancy)) +
  geom_point()
#plot life expectancy as a function of population size with x-axis to a log scale
ggplot(pop_liexp, aes(x= population, y=life_expectancy)) +
  geom_point()  +
  scale_x_continuous(trans='log2')
```
```{r}
#finding which years have missing data for infant mortality [using is.na function]
africadata$year[is.na(africadata$infant_mortality)]
#creating object with only 2000 year data
africadata2 <-africadata[africadata$year == 2000,]
#checking with str and summary
str(africadata2)
summary(africadata2)
#same plotting like before but only with year 2000 data
ggplot(africadata2, aes(x= infant_mortality, y = life_expectancy)) +
     geom_point()
ggplot(africadata2, aes(x= population, y=life_expectancy)) + 
     geom_point() + 
     scale_x_continuous(trans='log2')
```
```{r}
#fitting two linear regression models by regressing life expectancy on infant mortality and population size separately
fit1 <- lm(life_expectancy ~ infant_mortality, data = africadata2)
summary(fit1)
fit2 <- lm(life_expectancy ~ population, data = africadata2)
summary(fit2)
```
#fit1 has a significant p-value  indicates correlation between Infant Mortality and Life Expectancy

<br>

# Are any other variables signififcant predictors of life expectancy?

<br>

## First consider fertility 
```{r}
#still using data from the year 2000 for countries in Africa
#first consider fertility
#make a subset of data including fertility and life expectancy
africa_f_le <- africadata2 %>%
  select(fertility, life_expectancy)

#overview of fertility data structure
str(africa_f_le)

#summary of fertility data
summary(africa_f_le)

#to center the plot title
theme_update(plot.title = element_text(hjust = 0.5))

#plot life expectancy as a function of fertility
ggplot(data = africa_f_le, aes(x = fertility, y = life_expectancy)) +
  geom_point() +
  scale_size_manual(values = 2) +
  ggtitle("Life expectancy as a function of fertility") +
  labs(x = "Fertility", y = "Life Expectancy")

#create a linear model
fit3 <- lm(life_expectancy ~ fertility, data = africa_f_le)
tidy(fit3)

#plot of linear model
ggplot(africa_f_le, aes(x = fertility, y = life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  ggtitle("Life expectancy as a function of fertility with lm fit") +
  labs(x = "Fertility", y = "Life Expectancy")
```
<br>

## Conclusion: Fertility is a significant predictor of life expectancy in African countries in the year 2000 (t = -6.113, p < 0.001).

<br>

## Now consider GDP
```{r}
#make a subset of data including GDP and life expectancy
africa_gdp_le <- africadata2 %>%
  select(gdp, life_expectancy)

#overview of GDP data structure
str(africa_gdp_le)

#summary of GDP data
summary(africa_gdp_le)

#plot life expectancy as a function of GDP (on a log scale)
ggplot(data = africa_gdp_le, aes(x = gdp, y = life_expectancy)) +
  geom_point() +
  scale_size_manual(values = 2) +
  scale_x_continuous(trans='log2') +
  ggtitle("Life expectancy as a function of GDP") +
  labs(x = "log(GDP)", y = "Life Expectancy")

#create a linear model
fit4 <- lm(life_expectancy ~ gdp, data = africa_gdp_le)
tidy(fit4)

#plot of lm fit
ggplot(africa_gdp_le, aes(x = gdp, y = life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
   scale_x_continuous(trans='log2') +
  ggtitle("Life expectancy as a function of GDP with lm fit") +
  labs(x = "log(GDP)", y = "Life Expectancy")
```
<br>

## Conclusion: GDP is a significant predictor of life expectancy in African countries in the year 2000 (t = 2.516, p = 0.0152).