---
title: "Tidy Tuesday"
output:
  word_document:
    toc: no
  html_document:
    toc: no
---


# First install packages and load libraries
```{r}
library('readr')
library('readr')
library('ggplot2')
library("dplyr")
library("stringr")
library("scales")
```


# Read all the papers manually from the source
```{r}
papers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/papers.csv')
```
```{r}
glimpse(papers)
```
```{r}
paper_authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_authors.csv')
```

```{r}
glimpse(paper_authors)
```

```{r}
authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/authors.csv')
```

```{r}
glimpse(authors)
```

```{r}
paper_programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_programs.csv')
```
```{r}
glimpse(paper_programs)
```

```{r}
programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/programs.csv')
```

```{r}
glimpse(programs)
```
```


#View each of them and then write each of them as a csv file



```{r}
write_csv(x = papers, "Data/papers.csv")
write_csv(x = authors, "data/authors.csv")
write_csv(x = programs, "data/programs.csv")
write_csv(x = paper_authors, "data/paper_authors.csv")
write_csv(x = paper_programs, "data/paper_programs.csv")
```


#Clean the data and joined them [left joining]
```{r}
joined <- left_join(papers, paper_authors) %>% 
    left_join(authors) %>% 
     left_join(paper_programs) %>% 
     left_join(programs)%>% 
     mutate(
         catalogue_group = str_sub(paper, 1, 1),
         catalogue_group = case_when(
             catalogue_group == "h" ~ "Historical",
             catalogue_group == "t" ~ "Technical",
             catalogue_group == "w" ~ "General"
         ),
         .after = paper
     ) 
```
View(joined)


# See how many distinct values are there for each category
```{r}
sapply(joined, n_distinct)
```

#Let's see how many papers publishes each year throughout the 1973-2001 time
```{r}
ggplot(joined, aes(x=year, fill=catalogue_group)) + geom_bar() + labs(title= "Paper Count by year and cat_group between 1973 to 2001") + xlab("Year") + ylab("Count")
```

#It's quite clear that the majority is "general" type. So, let's do something else. What about finding number of papers according to program category? Lets do it
```{r}
ggplot(joined, aes(x=program_category)) +
             geom_bar(fill = "red") + labs(title= "Paper number by Program Category between 1973 to 2021") + xlab("Program Category") + ylab("Number")
```

#What about numbe rof papper for each year?             
```{r}
ggplot(joined, aes(x=year)) +
                             geom_bar(fill = "red") + labs(title= "Paper number by year between 1973 to 2021") + xlab("year") + ylab("Number")
```
#What about number of paper for each year categorized by program name also?
```{r}
ggplot(joined, aes(x=year, fill=program)) + geom_bar() + labs(title= "Paper number by year and program between 1973 to 2001") + xlab("Year") + ylab("Number")
```

#Most number of papers were in 2000. Let's look deeper in that
```{r}
data_2000 <- joined %>% filter(year ==2000)
```

#View the 2000 data
```{r}
glimpse(data_2000)
```

```{r}
#Let's see how many papers from each program got published in each month of 2000?
ggplot(joined, aes(x=month, fill=program_category)) + geom_bar() + labs(title= "Paper Count by month and program category in 2000") + xlab("Month") + ylab("Count")
```

```{r}
#We got a problem here. The x axis (month number is not integer. Let's fix it.

ggp <- ggplot(joined, aes(x=month, fill=program_category)) + geom_bar() + labs(title= "Paper Count by month and program category in 2000") + xlab("Month") + ylab("Count")

#fixed
ggp + scale_x_continuous(breaks = pretty_breaks())

```